---
title: "Examples"
---

This page provides examples of how to load and visualize the hackathon datasets using R.

## Setup

First, load the necessary libraries:

```{r}
#| message: false
library(sf)
library(tidyverse)
library(tmap)
```

## Loading Data

We will load the Super Data Zones (SDZ) and the filtered Origin-Destination (OD) data.

```{r}
#| message: false
# Ensure the data directory exists and contains the files
zones = read_sf("data/zones_sdz.gpkg")
od = read_csv("data/od_ni_open_filtered.csv")
```

## Visualizing Zones

Let's take a look at the Super Data Zones:

```{r}
qtm(zones)
```

## Analysing Commuting Flows

We can join the OD data to the zones to see which areas are the most popular destinations for commuters.

```{r}
# Aggregate total commuters by destination zone
dest_counts = od |> 
  group_by(workplace_area_code) |> 
  summarise(total_commuters = sum(count, na.rm = TRUE))

# Join back to zones
zones_joined = zones |> 
  left_join(dest_counts, by = c("sdz2021_cd" = "workplace_area_code"))

# Map the results
tm_shape(zones_joined) +
  tm_polygons("total_commuters", 
              style = "quantile", 
              title = "Total Commuters (Destination)",
              palette = "Viridis") +
  tm_layout(main.title = "Commuting Destinations in NI",
            frame = FALSE)
```

## Route Networks

Finally, let's visualize a snippet of the fastest route network.

```{r}
#| message: false
# Loading a subset for faster rendering
route_network = read_sf("data/route_network_fastest.gpkg")
qtm(head(route_network, 1000))
```
